
\begin{frame}[t]{Delivery strategy : Instant-ish }

\[ \begin{aligned}
\textrm{SURBs}\quad & \textrm{Rodger} \rightsquigarrow \textrm{Sam} \\
\textrm{Messages}\quad & \textrm{Sam} \to \cdots \to \textrm{Cross} \to \cdots \to R_k \to \textrm{Rodger}
\end{aligned} \]

Issues :
\begin{itemize}
\item Sam needs a {\em very recent} SURB from Rodger.  
\item $R_k$ should hold messages if Rodger went offline, but \\ 
 \hspace*{2pt} Rodger's ``Guard'' $R_k$ knows their IP. \\
\end{itemize}

% \pause

Solutions : 
\begin{itemize}
\item Rodger asks Sam not to use his SURBs for long. \\
\item Rodger sends SURBs to $R_k$ when they connect later.
\end{itemize}

\bigskip
Are these measures sufficient?  No. \\
 \hspace*{2pt} Rodger might have too many guard nodes.

\end{frame}


\begin{frame}[t]{Delivery strategy : Checked mailbox}
\[ \begin{aligned}
\textrm{Long SURBs}\quad & \textrm{Rodger} \rightsquigarrow \textrm{Sam} \\
\textrm{Messages}\quad & \textrm{Sam} \to \cdots \to \textrm{Cross}
  \to \cdots \to R_k \to \textrm{Mailbox} \\
\textrm{Short SURBs}\quad & \textrm{Rodger} \to \cdots \to \textrm{Mailbox} \\
\textrm{Messages}\quad & \textrm{Mailbox} \to \cdots \to R_k \to \textrm{Rodger} \\
\textrm{Deletion}\quad & \textrm{Rodger} \rightsquigarrow \textrm{Mailbox} 
\end{aligned} \]

\smallskip

We win lovely delivery properties from the cross-over point \\
 \hspace*{2pt} since Sam cannot flood, hack, etc. Rodger's mailbox server.

\smallskip

% Yet again :
\begin{itemize}
\item Sam needs a {\em less recent} and {\em longer lived} SURBs. % from Rodger.  
\item $R_k$ still holds messages if Rodger goes offline, but \\
 \hspace*{2pt} SURBs using $R_k$ are {\em short lived}. % and \\
% \hspace*{2pt} maybe Rodger confirms deletion on Mailbox.
\end{itemize}

% \pause

\medskip
Is this enough?  No.  \\
 \hspace*{2pt} No SURB can live forever!

\end{frame}


\begin{frame}[t]{Delivery strategy : Mailbox gateway}

We know good schemes for delivery credentials, so...

\[ \begin{aligned}
\textrm{Credential}\quad & \textrm{Rodger} \rightsquigarrow \textrm{Sam} \\
\textrm{Long SURBs}\quad & \textrm{Rodger} \to \cdots \to \textrm{Gateway} \\
\textrm{Messages}\quad & \textrm{Sam} \to \cdots \to \textrm{Gateway} \\
\textrm{Messages}\quad & \textrm{Gateway} \to \cdots \to R_k \to \textrm{Mailbox} \\
 \cdots
\end{aligned} \]

\bigskip
Is this one strategy to rule them all?  No. \\
 \hspace*{2pt} Sam can easily attack the gateway. \\
 \hspace*{2pt} And our instant-ish scheme is much faster.

\end{frame}
\begin{comment}
like 
\begin{itemize}
\item Single-use HMAC tokens to limit Sam's message sending, 
\item BBS or VLR group signatures for unlimited messages,
\item or nothing to allow a public contact point.
\end{itemize}
\end{comment}


