\documentclass[twoside,letterpaper]{sig-alternate}
\usepackage[margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage{eurosym}
\usepackage{tikz}
%\usepackage{listings}
%\usepackage{graphicx}
%\usepackage{wrapfig}
%\usepackage{caption}
%\usepackage{subcaption}

\usetikzlibrary{cd}
%\usetikzlibrary{shapes,arrows}
%\usetikzlibrary{positioning}
%\usetikzlibrary{calc}

\def\mathcomma{,}
\def\mathperiod{.}

\def\mathcomma{}
\def\mathperiod{}

\title{Identity-based encryption for meta-data protection when first establishing contacts}
% \subtitle{}
\author{Jeffrey Burdges, Florian Dold, Christian Grothoff, Neal Wa{\bf r}field}
\date{\today}

\begin{document}
\maketitle


\begin{abstract}
We argue that a limited use of identity-base encryption would improve
the security and confidentiality of some user-friendly end-to-end
encrypted messaging protocols.
\end{abstract}


\section{Introduction}

% \subsection{Setting}

A secure end-to-end encrypted channel requires that endpoints
authenticate one another, usually by verifying the fingerprint of
one another's public key material.
Assistance in verifying this material therefore becomes an important
feature. % of encryption protocols.

At the same time, the mechanism for obtaining key material tend to
leak meta-data about user's social graph, which can be harmful to
life and limb~\cite{skynet}.

Available decentralized solutions like the Web-of-Trust~\cite{wot}
leak social graph data since the key signing graph is public and
key lookups use a handful of key servers.  GNS provides a model
resembling the web-of-trust that leaks less profusely, but offers
no protection against confirmation attacks~\cite{gns}.
Both these schemes face deployment and adoption challenges.

Instead, there is wave of newer more user-friendly encrypted messaging
applications that distribute public key information using a centralized
trusted authorities, or sometimes trust-on-first-use (TOFU).

In Signal~\cite{TextSecure} for example, there is a central key server
from which clients obtain a pre-key that consists of a long-term key
and a preliminary Axolotl ratchet state.
This long term key provides a fingerprint that users could verify
out-of-band, or verbally, for authentication.

There are two fundamental problems with using centralized key servers
in this way.
First, users tend not to actually verify fingerprints, so
key servers could be coerced into substituting a false key.
% Signal itself actively encourages users to authenticate fingerprints,
% but proprietary efforts tend not to show fingerprints by default.
Second, the authority learns who resolved which names.

%Signal's Axolotl ratchet has impressive forward-secrecy properties
%that strengthen any protocol parts that depend upon TOFU, and
%reduces the sheer diversity of attacks on systems like PGP.
%Of course, forward-secrecy cannot replace authentication.
%% Yet, Signal provides deniability for example.

\smallskip
% \subsection{Goals}

We propose that, systems like Signal that already rely on an authority,
should avoid exposing meta data to the authority by using
identity-based encryption (IBE) when first establishing a contact.
In the first few messaged, users exchange long-term fingerprint key and
ratchet state, which are locally generated, and not identity-based.

In essence, we employ IBE only for the weak form of authentication provided
by the trusted authentication server,
while full authentication still requires that users verify fingerprints.
In doing so, we prevent meta data leakage to the authority, and reduce
its attack surface for manipulating pre-keys.

The same scheme also naturally strengthens designs based on TOFU:
an authority can be introduced that provides IBE-based keys to
whoever controls a particular network address.  This defeats an
adversary who has control over the link between the client and the
server, but not the link between the server and the authority.
Such an adversary would be successful against pure TOFU, but fail
against TOFU plus certification by an authority.  Again IBE ensures
that the authority only learns about the set of participants, but
not their relationships with each other.

Our scheme remains fundamentally vulnerable to an attacker who can both
extract the recipient's private key from the authority, and
attack the transport during the initial exchange.
In particular, we loose if the authority's servers provide the transport
layer as well.  Naturally, having the authorty in control of the
transport layer also leaks the social graph, so this is not desireable
for multiple reasons.

To avoid relying on the authority's servers to resolve addresses and
strictly avoid leaking meta data on the transport layer, we need an
anonymity preserving transport layer where recipients are determined
using an identity.  This could be implemented by adapting a system
like Tor~\cite{tor} to support an identity-based naming scheme
for hidden services~\cite{hidden}.

Finally, we propose to decentralize the IBE scheme's private key
issuing authority by allowing contact initiators to expect the
responder to obtain private keys from a set of authorities selected by
the initiator.  This way, an adversary would have to compromise
potentially many paths in the network or multiple authorities.
Consequently, the attack is only likely to succeed if the adversary
controls most network links to the victim, and even then manual
fingerprint verification might still reveal the attack.


\section{Ratchet spin-up protocol}

\def\E{\mathbb{E}}
\def\F{\mathbb{F}}
\def\Z{\mathbb{Z}}
\def\ID{\mathtt{ID}}
\def\rk{\mathtt{rk}}
\def\ck{\mathtt{ck}}
\def\hk{\mathtt{mk}}
\def\mk{\mathtt{mk}}

\subsection{Around pairing}

Almost all identity-based scheme employ pairing based cryptography,
or else relate to RSA.  We use pairing to avoid the extra homomorphic
properties surounding RSA. 

In general in pairing based cryptography, and
 in the Boneh–Franklin IBE scheme ~\cite{BF-IBE} in particular,
we have an elliptic curve $E_1$ over a field $\F_1$,
 along with an element $P_1 \in E_1$ of order $q$.
In addition, we have a group $E_2$ of order $q$ and
 a pairing $e : E_1 \times E1 \to \mathbb{E}_2$.

As a rule, pairing friendly curves are large, slow, and
 relatively weak cryptographically,
 actually no stronger than RSA~\cite{??paring_weakness??}.
It follows that, after this initial contact request, we should
transition the ratchet to a faster higher security curve $\mathbb{E}$,
such as curve25519~\cite{DJB-Curve25519}
 or Ed448-Goldilocks~\cite{Ed448-Goldilocks}.
Let $P$ denote the distinguished generator of $\mathbb{E}$.

\subsection{Identity-based scheme}
% \medskip

Along with a long-term fingerprint key,
a Signal pre-key contains the initial elliptic curve point required
by the Axolotl ratchet, meaning
 the first message can operate the ratchet as usual.
We shall instead derive the initial root key from a identity-base key
exchange, which we prefer over a public key system,
 like Sakai–Kasahara scheme (SAKKE),
for simplicity when integrating with the ratchet.
Indeed, we even choose to employ the chain key machinery of Axolotl,
 in case we need to send multiple messages.

We should, however, choose a key exchange that does not leak the recipient's
identity, which many IBE schemes do~\cite{AnonIBE}.
A good choice would be the Boneh–Franklin IBE scheme~\cite{BF-IBE}
(see \cite[??]{BoyenMIBS}).

We employ essentially the usual Boneh–Franklin IBE setup phase:
Our private key issuing authority
 selects a private key $s \in \F^\times$,
 derives the public key $K = s P_1$, and
 selects two public hash functions
  $H_1 : \{0,1\}^* \to \E_1$ and $H_1 : E_2 \to \{0,1\}^n \to \E_1$.
Amongst these values, both $s$ and $H_1$ should together be rotated
 periodically, such as daily or weekly,
while allowing for clock skew, and possibly for delayed messages.
$H_1$ could for example be seeded by $K$.
% Is this actually important for $d_{\ID || F}$?

We adjust the usual Boneh–Franklin IBE extraction phase only slightly:
At some point, user claims ownership of an identity $\ID$, thereby
 associating with it a public key $F = f P \in \E$.
In practice, identities would be phone numbers or e-mail addresses,
and proving possession is done by demonstrating that
one is reachable at the respective address.
We now have the usual private and public identity-based keys, given by
 $d_\ID = s H_1(\ID)$ and $g_{\ID_R} = e(H_1(\ID_R),K)$, respectively.
Our private key issuing authority issues the two private keys
 $d_\ID = s H_1(\ID)$ and $d_{\ID,F} = s H_1(\ID || F)$
to anyone who can prove possession of the private key $f$ for $F$
 by signing a request. 

\subsection{Axolotl tweaks} % motivation/outline
% \medskip

We first outline our tweaks to the Axolotl ratchet before
 spelling out the protocol in more detail.

We cannot transform the initiator's key pair for
 our pairing friendly curve system $(\E_1,P_1,\E_2,e,\ldots)$
into a key pair on $\E$~\cite{??no_homomorphism??}.
As a result, any message sent on this initial Axolotl chain must include
both our first Axolotl public key $B = b P \in \mathbb{E}$,
 along with our long-term fingerprint key $F \in \mathbb{E}$.

We mentioned above using the chain key machinery of Axolotl even from
the first message.  There are risks in this decision because the
authority can decrypt these messages.  It appears sufficient that
contact request messages either disallow user content all together,
perhaps a desirable feature for preventing SPAM in any case, or that
perhaps users merely be warned not to speak freely.

As usual in Axolotl, we must hold fixed the initiator's public side of
the identity-base key exchange throughout the lifetime of this initial
chain key.  Interestingly, we need not necessarily retain the
initiator's private key for the identity-base key exchange, and
 the responder need only run the key exchange once as well.
We do retain the private key $b$ for $B$ until the responder uses it.

As a final tweak to Axolotl, we should incorporate the long-term
fingerprint key by using triple Diffie-Hellman during early rounds.
We still lack confidence in a user's fingerprint until
we have seen it used however.
We could speed this up by using two long-term fingerprint keys,
an initiator one in the pairing system, and
 a responder key for $\mathbb{E}$,
with the fingerprint being the hash of both.

\subsection{Axolotl chain one}
% \medskip

We begin a fresh contact with all Axolotl key material undefined, 
 focusing our discussion first on ratchet cold start situation.
We shall also describe adjustments that allow identity-based
information to be adjoined to a running ratchet,
as this buys us identity-based authentication for
 the symetric half of our key exchange.

\smallskip

An initiator knows their responder's identity $\ID_R$, so they may
compute the responder's public key as $g_{\ID_R} = e(H_1(\ID_R),K)$.
To setup a contact, the initiator first chooses a random $r \in \Z/q\Z$
to compute the ephemeral public key $U = r P_1$, 
 compute $\rho_0 = H_2(r g_{\ID_R})$,
 assign the first root key $\rk = \rho_0$, and 
 compute the outgoing chain key $\ck_I = H(\mathrm{``Chain''} || \rk)$.
In addition, they choose a scalar $b$ for $\E$,
 and then save $(\rk,\ck_I,r,b)$.

In this first round, we would love to communicate $F_I$ or $\ID_I$ where
 $F_I$ is the initiator's long-term fingerprint key, and
 $\ID_I$ is the initiators identity.
We cannot however because this would expose them to
 a collusion between the transport and private key authority.
If that were not a concern, then one might adjust the key exchange to
 run both halfs almost in parallel.

An initial message must communicate $B$ where
 $B = b P$ is the initiator's first regular Axolotl public-key,
along with additional information that poses no risk if exposed.
So, to send the contact request, the initiator
sends $U$ along with $E_{\mk}(B,\ldots)$ where
 $\mk = H(\mathrm{``Message''}  || \ck_I)$ is the message key,
and replaces the saved $\ck_I$ with $H(\mathrm{``Iterate''} || \ck_I)$.

Of course, this sending process may be rerun with different
additional information, but it remain only as secure as $d_{\ID_R}$
and vulnerable to the authority.

\smallskip

To decode the contact request message, the respondent
first computes $\rho_0 = H_2( e(d_{\ID_R}, U) )$,
 assign the first root key $\rk = \rho_0$, and 
computes their incoming chain key $\ck_I = H(\mathrm{``Chain''} || \rk)$.
Our two computaitons of the shared secrets $\rho_0$ agree because
\begin{align*}
e(d_{\ID_R}, U) = e(s H_1(\ID_R), r P) \\
 \quad = r e(H_1(\ID_R), s P) = r g_{\ID_R}
\end{align*}
Next they compute the initiator's
 message key $\mk = H(\mathrm{``Message''}  || \ck_I)$
to attempt to decrypt the message body.
If this succeeds, the respondent learns and saves $(B,\ldots)$
 under the index $H(U)$.
If not, they may save the skipped message keys $\mk$ under
the index $H(U)$, and iterate $\ck_I$, as usual in Axolotl.

\smallskip % \subsection{Non-cold start}

If we run the above with an established ratchet, then we update
 the root key $\rk$ to be $H(\rk || x Y || \rho_0 )$ as opposed to
 just assigning it to be $\rho_0$.
In this expression, $Y$ is our contact's public point in $\E$, 
 making it either $B$, or the predecessor of $A$,
while $x$ is our private scalar,
 making it either $b$, or the predecessor of $a$.
Also, $B$ must appear in the header instead of the message.
In fact, $B$ could appear in the header anyway, 
 assuming the MAC protects it appropiately.

If header encryption is being used, then
 the header keys can be initalized or update with the root key.
$U$ cannot appear in the header during start up.

\subsection{Axolotl chain two}

We recall that $d_{\ID,F} = s H_1(\ID || F)$ and
 $g_{\ID,F} = e(H_1(\ID || F),K)$.

To accept the contact request, the respondent
first chooses a random scalar $a$ for $\E$ to
iterate to the second root key $\rk := H(\rk || a B)$ and
 compute their outgoing chain key $\ck_R = H(\mathrm{``Chain''} || \rk)$.
The respondent also computes $\rho_1 := H_2( e(d_{\ID_R,F_R}, U) )$
and saves $(\rk,\ck_I,\ck_R,B,a,\rho_1)$.

An accept message should communicate 
 respondent's first regular Axolotl public-key $A = a P$,
 the respondent's long-term fingerprint key $F_R$, and
 a MAC incorporating $\rho_1$.
So, to send the acceptance, the respondent
sends $H(B)$, $A$, and $E_{\mk}(F_R,\ldots)$ where again
 $\mk = H(\mathrm{``Message''}  || \ck_R)$ is the message key,
and replaces the saved $\ck_R$ with $H(\mathrm{``Iterate''} || \ck_R)$.

\smallskip

An initiator processes the acceptance by recognizing $H(B)$,
 thereby learning $A$.
At this point, they iterate to the second root key
 $\rk := H(\rk, b A )$, and
compute the incoming chain key $\ck_R = H(\mathrm{``Chain''} || \rk)$.
Next they compute the respondent's
 message key $\mk = H(\mathrm{``Message''}  || \ck_R)$
to attempt to decrypt the message body.
If this succeeds, the initiator learns and saves $(A,F_R,\ldots)$.
If not, they may save the skipped message keys $\mk$ under the index $H(U)$,,
and iterate $\ck_R$, as usual in Axolotl.
In addition, they compute $\rho_1 = H_2( u g_{\ID,F} )$,
 which again agrees with the first computation of $\rho_1$,
so they can verify the MAC  incorporating $\rho_1$.

\smallskip







At this point, we have exhausted our identity-base key exchange scheme,
 as both sides were hashed into the the root key.  In addition, we have
hashed in one exchange $a B = b A$ between strong ephemeral keys, and
one exchange $a F_I = f_I A$ between an ephemeral and
 the long-term fingerprint key $F_I = f_I P$.
It follows that both parties know one another's finger prints, but only
the initiator knows that the respondent saw the correct fingerprint key.
To address this with fewer round trips, we ask that both parties
now fast-forward the root key $\rk$ by replacing it with
 $\rk' := H(\rk f_R B) = H(\rk, b F_R )$.
Both parties may do so as soon as they have computed the respondent's
 first chain key $\ck_R$.
We thus allow Axolotl to run as usual from this point forward, while
ensuring that the initiator's next message conveys that the initiator
 has the correct long-term fingerprint key.


\section{Identity-based mixnet} % transport

... describe Tor hidden services with identity-based keys ...


\section{Analysis}

We observe that one principle weakness of our proposal is that pre-keys
offer a degree of ephemerality lacking in identity-based keys.

If for example a phone is stolen, then a Signal user can simply ask
for a new SIM card from their telephone provider, and install Signal
on their new phone.  On install, Signal replaces all their pre-keys
stored on the key server, invalidating the old ones, and restarts all
the ratchet states with their contacts, thereby making the old phone's
data worthless for impersonation.

In our identity-based scheme, the old phone retains the victims
identity-based private key $d_{\ID}$, which the thief could exploit for
either impersonation or eves dropping on initial contact request messages.
We could mitigate theft risks by rotating the authority's private key $s$,
perhaps daily or weekly, but this attack window remains larger than with Signal.

In this vein, an attacker who infiltrates the phone system could steal
the identity-based private key $d_{\ID}$ anytime they like.
Any such attacker could replace pre-keys and restart ratchets as well,
but this appears tougher to exploit.

This attack is easily addressed by adding a {\em date} to the
identity of the user in the IBE scheme.  Honest authorities would
only hand out private keys for the current date (modulo being
a tad permissive to account for some clock skew).  The date forces
the adversary to take control of the identity on the same day, as
the inclusion of the date changes the IBE key every day.

\smallskip


\smallskip

...

There is an interesting unknown key share attack on Signal that
even verifying fingerprints cannot address~\cite[\S4.2]{TextSecure}.
At least one proposal for fixing the Unknown Key Share attack
modifies the pre-keys to leak not only all new contacts, but
all currently known possible contacts~\cite[\S4.3]{TextSecure}!
% FIXME Answer: Read \S4.2 and \S4.3 in TextSecure if the above makes no sense.
As with many pre-key designs, this fix may damage the deniability
 of an early message by incorporating a signature.

We avert the unknown key share attack under the assumption that
an attacker cannot learn private IBE keys from the authority, and
 without causing a toxic metadata spill.
We also avoid sacrifice deniability during these initial couple steps of
the ratchet by instead incorporating key exchanges with long-term
fingerprint keys, much like triple Diffie-Hellman does.
We pay a price in delaying certainty of the fingerprint, but
 that is unavoidable in our context.


Alternatively, one could employ triple Diffie-Hellman inside the
Axolotl ratchet during two initial steps,
 thus sacrificing deniability during those steps,
 and delaying certainty of the fingerprint.



\section{Decentralization}

Signal could reduce the vulnerability of pre-keys by using
multiple pre-keys distributed to multiple organizations,
 possibly in different legal environment, but
doing so risks exposing user's meta-data to more parties.

In our identity-base scheme, there is no such risk from using
multiple key generation authorities in different jurisdictions.

... describe or cite multi-authority IBE ...


\section{Conclusions}

...


% \section*{Acknowledgements}
% This work benefits from the financial support of the Brittany Region
% (ARED 9178) and a grant from the Renewable Freedom Foundation.


%\newpage

\bibliographystyle{abbrv}
\bibliography{msg,ecc,ibe}

\end{document}


\section{}
